---
title: "imputing data for NCDB"
author: "Michelle.VanTieghem"
date: "10/8/2020"
output: html_document
---


```{r}

library(tidymodels)
library(tidyverse)
library(readxl)
library(mice)
```

```{r}
df = read.csv("../data/cleaned/NCDB_cleaned_all_cases.csv")
nrow(df)
```

select for variables we want to keep, 
and exclude cases missing outcome and impt varr
```{r}
df_select = df %>%
  select(AGE, tumor_size_bins_cm, lymph_vasc_invasion, metastasis) %>%
  filter(!is.na(metastasis) & !is.na(lymph_vasc_invasion))

nrow(df_select)
```


## check tumor size 
```{r}
summary(df_select$tumor_size_bins_cm)

df_select %>%
  filter(tumor_size_bins_cm == 0) %>%
  count()
```

tumor size colinear with other vars?
```{r}
ggplot(data = df_select, aes(x = lymph_vasc_invasion, y = tumor_size_bins_cm)) + 
  geom_jitter() 

```
examine missingness
```{r}
md.pattern(df_select)

```


## impute missing data
m = the numbe of imputed datasets, 5 is default 
method = lmm is predictive mean matching 
will only impute tumor size
```{r, warnings = F}
tempData = mice(df_select, m = 5, maxit = 50, meth = 'pmm', seed = 500)
summary(tempData)

```

check imputed data

```{r}
# red is imputed, blue is observed.
densityplot(tempData)
```


```{r}
# columns = imputation set 
# rows = cases with missing data
head(tempData$imp$tumor_size_bins_cm)

# save the indices for imputed data
index_imputed = rownames(tempData$imp$tumor_size_bins_cm)
```

Add imputed data 
```{r}
completeData = complete(tempData, 1)

# add marker for imputed variables 
completeData$imputed = NA
for (i in 1:nrow(completeData)){
 # print(i)
  if (i %in% index_imputed){
    completeData$imputed[i] = 1
  } else { 
    completeData$imputed[i] = 0}
}
  
  
summary(as.factor(completeData$imputed))

```

## check distributions of imputed data by metastasis
```{r}
ggplot(data =  completeData, aes(tumor_size_bins_cm,
                                 color = as.factor(imputed))) + 
  geom_density() +
  theme_bw() + facet_grid(~metastasis)

```

## inspect relationships between imputed variable and LVI
```{r}
table <- completeData %>%
  group_by(lymph_vasc_invasion) %>%
  summarize(means = mean(tumor_size_bins_cm), 
            sd = sd (tumor_size_bins_cm), 
            lower = means -sd, 
            upper = means + sd)

ggplot(data =  table, aes(x = as.factor(lymph_vasc_invasion),
                          y  = means))  +  
  geom_jitter(data = completeData, aes(x = as.factor(lymph_vasc_invasion),
                                       y = tumor_size_bins_cm),
              alpha = 0.2) +
  geom_errorbar(color = "dark red", size = 1, width = 0.2,
                aes(ymin = lower, ymax = upper)) +
    geom_point(color = "dark red", size = 3) + theme_bw() +
  ylab("Tumor size")

  
```


```{r}

write.csv(completeData, file = "../data/processed/NCDB_MICE_imputed.csv", row.names = FALSE)
```